Artin Group Presentations of Reflection Groups

### 2014-07-16 Paper.tex
small changes

%\usepackage[all]{xy}
\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{rem}[thm]{Remark}
\newcommand\fri{s_4^{-1}}
\newcommand\fr{s_4}
\newcommand\tri{s_3^{-1}}
\newcommand\tr{s_3}
\newcommand\twi{s_2^{-1}}
\newcommand\tw{s_2}
\newcommand\oni{s_1^{-1}}
\newcommand\on{s_1}
\newcommand{\V}{$V(\Gamma)$}

@@

\title{Artin Group Presentations of Reflection Groups}
\author{Jacob Haley, David Hemminger, Aaron Landesman, Hailee Peck}
\address{you can put addresses for authors\\
here \\
}
\email{..., hpeck@millikin.edu}
\date{\today}
\keywords{Artin group, cluster algebra, diagram representations}

@@

\section{Introduction \& Motivation}

In \cite{FZ02}, Fomin and Zelevinsky first introduced the concept of cluster algebras. This introductory paper focused on structural features of cluster algebras, specifically that, in a given cluster, when any cluster variable is viewed as a rational function in the variables of the given cluster, this cluster variable is a Laurent polynomial. Fomin and Zelevinsky initially defined this type of commutative ring in order to make further strides in the areas of representation theory, Lie theory, and total positivity. Since then, the study of cluster algebras has provided a motivation for applications in various other areas of mathematics, including quiver representations. Of particular interest were \textit{finite-type} cluster algebras; that is, cluster algebras whose variables are generated through mutation on a finite number of \textit{seeds}. In the 2003 sequel to their introductory paper (\cite{FZ03}), Fomin and Zelevinsky introduce the concept of \textit{mutation equivalence} between diagrams, proving that a connected graph is mutation equivalent to an oriented Dynkin diagram if and only if all mutation equivalent graphs have edge weights not exceeding 3. In particular, this proves that finite-type cluster algebras can be classified by Dynkin diagrams. \\
\indent Barot and Marsh extended Fomin and Zelevinsky's results in \cite{BM13}, providing a presentation of the reflection group associated to a Dynkin diagram with generators that correspond to elements of a companion basis associated to a seed of a finite-type cluster algebra. They also proved that this group presentation is invariant up to isomorphism under the mutation equivalence relation. That is, given a diagram $\Gamma$ and a diagram mutation equivalent to $\Gamma$, denoted $\Gamma^{\prime} = \mu_k(\Gamma)$, they proved that $W_{\Gamma} \cong W_{\Gamma^{\prime}}$, where $W_{\Gamma}$ and $W_{\Gamma^{\prime}}$ are the group representations corresponding to $\Gamma$ and $\Gamma^{\prime}$, respectively. \\
\indent Section 2 provides the necessary definitions and fundamental results from \cite{BM13} to provide motivation for our own results. For further definitions and references on the topic, we refer the reader to \cite{FZ02}. Section 3  will review theory from \cite{FZ02}, \cite{FZ03} as well as review the classifications (from \cite{BM13}) of mutations of diagrams and their oriented chordless cycles. In this section we also define the appropriate relations for our Artin group presentations. Section 4 specifies a relation for each case of the chordless cycles which suffices to show all the other relations in the group presentation hold. Finally, section 5 will provide the proof that the Artin group defined for a diagram $\Gamma$ is invariant up to isomorphism under mutations of $\Gamma$. \\

@@

A \textit{cluster algebra} is a commutative ring with unity but without zero-divisors, and it can be generated by a set of generators called \textit{cluster variables} that satisfy certain exchange relations. Following the style of \cite{FZ02} and \cite{BM13}, we will define cluster algebras in terms of \textit{skew-symmetrisable} matrices (that is, a matrix $B$ such that there exists a diagonal matrix $D$ of the same size with $D_{ii} >0$ such that $DB$ is skew-symmetric). Let $\mathbb{F} = \mathbb{Q}(u_1, u_2, \ldots, u_n)$ be the field of rational functions in $n$ indeterminates over $\mathbb{Q}$. We will define an \textit{intial seed} for the cluster algebra to be a fixed pair $(\textbf{x}, B)$, where $\textbf{x} = \left{x_1, \ldots, x_n\right}$ is a free generating set of $\mathbb{F}$ and $B$ is an $n \times n$ skew-symmetric matrix. Define $x_k^{\prime} \in \mathbb{F}$ by the \textit{exchange relation}
\begin{displaymath}
x_k^{\prime}x_k = \prod_{B_{ik} > 0}{x_i^{B_{ik}}} + \prod_{B_{ik} < 0}{x_i^{--B_{ik}}}
\end{displaymath}
Then, given an initial seed $(\textbf{x}, B)$ and $k \in {1,2,\ldots,n}$, we can define a \textit{mutation} of the seed at $k$, denoted $\mu_k(\textbf{x}, B) = (\textbf{x^{\prime}}, B^{\prime})$ where:
\begin{displaymath}
B_{ij}^{\prime} = \begin{cases} -- B_{ij} & \mbox{ if } i = k \mbox{ or } j = k;\\
																B_{ij} + \frac{|B_{ik}|B_{kj} + B_{ik}|B_{kj}|}{2} & \mbox{ otherwise. }\\
									\end{cases}
\end{displaymath}
and $\textbf{x^{\prime}} = {x_1, x_2, \ldots, x_{k-1}, x_k^{\prime}, x_{k+1}, \ldots, x_n}$. 
Such a mutation or a sequence of such mutations generate \textit{seeds} which in turn generate all cluster variables in that, for each $\textbf{x} = {x_1, \ldots, x_n}$ corresponding to a seed of the cluster algebra, the entries $x_i$ are the cluster variables. \\

@@

Just as we can define mutations of the seeds of a cluster variable, we can also define mutations of a diagram associated to a cluster algebra by the following set of rules:
\begin{proposition}\cite{BM13}[Proposition 1.4]
Let $B$ be a 2-finite skew-symmetrisable matrix. Then $\Gamma(\mu_k(B))$ is uniquely determined by $\Gamma(B)$ as follows: 
\begin{itemize}
\item Reverse the orientations of all edges in $\Gamma(B)$ incident with $k$ (leaving the weights unchanged)
\item For any path in $\Gamma(B)$ of form $i \xrightarrow{a} k \xrightarrow{b} j$ (i.e. with $a,b$ positive), let $c$ be the weight on the edge $j \rightarrow i$, taken to be zero if there is no such arrow. Let $c'$ be determined by $c'\geq 0$ and 
$c+c' =$ max$(a,b)$. 
Then $\Gamma(B)$ changes in a predetermined way, taking the case $c' = 0$ to mean no arrow between $i$ and $j$.
\end{itemize}
\end{proposition}

@@

Then, they define $W(\Gamma)$ to be the group generated by $s_i$, for $i \in \V$, under the following relations. Note that $e$ will denote the identity element of $W(\Gamma)$. 
\begin{enumerate}
\item $s_i^2 = e$ for all $i$;
\item $(s_is_j)^{m_{ij}} = e$ for all $i \neq j$;
\item For any chordless cycle $C$ in $\Gamma$, where
\begin{displaymath}
C = i_0 \rightarrow i_1 \rightarrow \cdots \rightarrow i_{d-1} \rightarrow i_0
\end{displaymath}
and all of the weights $w_k$ are $1$ or $w_0 = 2$, we have
\begin{displaymath}
(s_{i_0}s_{i_1}\cdots s_{i_{d-2}}s_{i_{d-1}}s_{i_{d-2}}\cdots s_{i_1})^2 =e.
\end{displaymath}

Using this group presentation, Barot and Marsh state the following result:
\begin{theorem}\cite{BM13}[Theorem A]
Let $\Gamma$ be the diagram associated to a seed in a cluster algebra of finite type. Then $W(\Gamma)$ is isomorphic to the corresponding reflection group.
\end{theorem}

@@

\begin{definition}
Let $W_{\Gamma}$ be the group with generators $s_i, i = 1,2,\ldots, n$, subject to the following relations:
\begin{itemize}
\item{(R1)} $s_i^2 = e$ for all $i$
\item{(R2)} $\left(s_is_j\right)^{m_{ij}} = e$ for all $i \neq j$
\end{itemize}
Furthermore, for a chordless cycle $C : i_0 \rightarrow i_1 \rightarrow \cdots \rightarrow i_{d-1} \rightarrow i_0$ and for $a = 0,1,2,\ldots, d-1$, define \textbf{$r\left(i_a, i_{a+1}\right) = s_{i_a}s_{i_{a+1}} \cdots s_{i_{a+d-1}}s_{i_{a+d-2}} \cdots s_{i_{a+1}}$}.\\

\vspace{0.1cm}
Then we have the following relations:
\begin{itemize}
\item{(R3)(a)} If all the weights in the edges of $C$ are $1$, then $r(i_a, i_{a+1})^2 = e$
\item{(R3)(b)} If $C$ has some edges of weight $2$, then $r(i_a, i_{a+1})^k = e$ where $k = 4-w_a$ and $w_a$ is the weight of the edge $i_a\--\ i_{a-1}$
\end{itemize}
\end{definition}

Defining the group $W_{\Gamma}$ with relations as shown above allows them to prove certain characteristics about the interaction between the relations in this group for the chordless cycles underlying the diagrams in question. In particular, they prove the following result, which we include as our main goal was to define analogous relations for an Artin group that would satisfy the same result.
\begin{theorem}\cite{BM13}[Theorem 5.4a]
Let $\Gamma$ be a diagram of finite type and $\Gamma^{\prime} = \mu_k(\Gamma)$ the mutation of $\Gamma$ at vertex $k$. Then $W_{\Gamma} \cong W_{\Gamma^{\prime}}$.
\end{theorem}

@@

\section{Main Results}

>>>>

%\usepackage[all]{xy}
\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{rem}[thm]{Remark}
\newcommand\fri{s_4^{-1}}
\newcommand\fr{s_4}
\newcommand\tri{s_3^{-1}}
\newcommand\tr{s_3}
\newcommand\twi{s_2^{-1}}
\newcommand\tw{s_2}
\newcommand\oni{s_1^{-1}}
\newcommand\on{s_1}
\newcommand\V{V(\Gamma)}

@@

\title{Artin Group Presentations of Reflection Groups}
\author{Jacob Haley, David Hemminger, Aaron Landesman, Hailee Peck}
\address{you can put addresses for authors\\
here \\
}
\email{...,david.hemminger@duke.edu, hpeck@millikin.edu}
\date{\today}
\keywords{Artin group, cluster algebra, diagram representations}

@@

\section{Introduction \& Motivation}\todo{let's add the statement of the main theorem to the introduction - David}

\todo{use section labels rather than explicit numbers}
In \cite{FZ02}, Fomin and Zelevinsky first introduced the concept of cluster algebras. This introductory paper focused on structural features of cluster algebras, specifically that, in a given cluster, when any cluster variable is viewed as a rational function in the variables of the given cluster, this cluster variable is a Laurent polynomial. Fomin and Zelevinsky initially defined this type of commutative ring in order to make further strides in the areas of representation theory, Lie theory, and total positivity. Since then, the study of cluster algebras has provided a motivation for applications in various other areas of mathematics, including quiver representations. Of particular interest were \textit{finite-type} cluster algebras; that is, cluster algebras whose variables are generated through mutation on a finite number of \textit{seeds}. In the 2003 sequel to their introductory paper (\cite{FZ03}), Fomin and Zelevinsky introduce the concept of \textit{mutation equivalence} between diagrams, proving that a connected graph is mutation equivalent to an oriented Dynkin diagram if and only if all mutation equivalent graphs have edge weights not exceeding 3. In particular, this proves that finite-type cluster algebras can be classified by Dynkin diagrams. \\
\indent Barot and Marsh extended Fomin and Zelevinsky's results in \cite{BM13}, providing a presentation of the reflection group associated to a Dynkin diagram with generators that correspond to elements of a companion basis associated to a seed of a finite-type cluster algebra. They also proved that this group presentation is invariant up to isomorphism under the mutation equivalence relation. That is, given a diagram $\Gamma$ and a diagram mutation equivalent to $\Gamma$, denoted $\Gamma^{\prime} = \mu_k(\Gamma)$, they proved that $W_{\Gamma} \cong W_{\Gamma^{\prime}}$, where $W_{\Gamma}$ and $W_{\Gamma^{\prime}}$ are the group representations corresponding to $\Gamma$ and $\Gamma^{\prime}$, respectively. \\
\indent Section 2 provides the necessary definitions and fundamental results from \cite{BM13} to provide motivation for our own results. For further definitions and references on the topic, we refer the reader to \cite{FZ02}. Section 3  will review theory from \cite{FZ02}, \cite{FZ03} as well as review the classifications (from \cite{BM13}) of mutations of diagrams and their oriented chordless cycles. In this section we also define the appropriate relations for our Artin group presentations. Section 4 specifies a relation for each case of the chordless cycles which suffices to show all the other relations in the group presentation hold. Finally, section \todo{for example}\ref{sec:proof_of_main} will provide the proof that the Artin group defined for a diagram $\Gamma$ is invariant up to isomorphism under mutations of $\Gamma$. \\

@@

A \textit{cluster algebra} is a commutative ring with unity but without zero-divisors, and it can be generated by a set of generators called \textit{cluster variables} that satisfy certain exchange relations. Following the style of \cite{FZ02} and \cite{BM13}, we will define cluster algebras in terms of \textit{skew-symmetrisable} matrices (that is, a matrix $B$ such that there exists a diagonal matrix $D$ of the same size with $D_{ii} >0$ such that $DB$ is skew-symmetric). Let $\mathbb{F} = \mathbb{Q}(u_1, u_2, \ldots, u_n)$ be the field of rational functions in $n$ indeterminates over $\mathbb{Q}$. We will define an \textit{intial seed} for the cluster algebra to be a fixed pair $(\textbf{x}, B)$, where $\textbf{x} = \{x_1, \ldots, x_n\}$ is a free generating set of $\mathbb{F}$ and $B$ is an $n \times n$ skew-symmetric matrix. Define $x_k^{\prime} \in \mathbb{F}$ by the \textit{exchange relation}
\begin{displaymath}
x_k^{\prime}x_k = \prod_{B_{ik} > 0}{x_i^{B_{ik}}} + \prod_{B_{ik} < 0}{x_i^{--B_{ik}}}
\end{displaymath}
Then, given an initial seed $(\textbf{x}, B)$ and $k \in {1,2,\ldots,n}$, we can define a \textit{mutation} of the seed at $k$, denoted $\mu_k(\boldsymbol{x}, B) = (\textbf{x}^{\prime}, B^{\prime})$ where:
\begin{displaymath}
B_{ij}^{\prime} = \begin{cases} -- B_{ij} & \mbox{ if } i = k \mbox{ or } j = k;\\
																B_{ij} + \frac{|B_{ik}|B_{kj} + B_{ik}|B_{kj}|}{2} & \mbox{ otherwise. }\\
									\end{cases}
\end{displaymath}
and $\textbf{x}^{\prime} = {x_1, x_2, \ldots, x_{k-1}, x_k^{\prime}, x_{k+1}, \ldots, x_n}$. 
Such a mutation or a sequence of such mutations generate \textit{seeds} which in turn generate all cluster variables in that, for each $\textbf{x} = {x_1, \ldots, x_n}$ corresponding to a seed of the cluster algebra, the entries $x_i$ are the cluster variables. \\

@@

Just as we can define mutations of the seeds of a cluster variable, we can also define mutations of a diagram associated to a cluster algebra by the following set of rules:
\begin{prop}\cite{BM13}[Proposition 1.4]
Let $B$ be a 2-finite skew-symmetrisable matrix. Then $\Gamma(\mu_k(B))$ is uniquely determined by $\Gamma(B)$ as follows: 
\begin{itemize}
\item Reverse the orientations of all edges in $\Gamma(B)$ incident with $k$ (leaving the weights unchanged)
\item For any path in $\Gamma(B)$ of form $i \xrightarrow{a} k \xrightarrow{b} j$ (i.e. with $a,b$ positive), let $c$ be the weight on the edge $j \rightarrow i$, taken to be zero if there is no such arrow. Let $c'$ be determined by $c'\geq 0$ and 
$c+c' =$ max$(a,b)$. 
Then $\Gamma(B)$ changes in a predetermined way, taking the case $c' = 0$ to mean no arrow between $i$ and $j$.
\end{itemize}
\end{prop}

@@

Then, they define $W(\Gamma)$ to be the group generated by $s_i$, for $i \in \V$, under the following relations. Note that $e$ will denote the identity element of $W(\Gamma)$. 
\begin{enumerate}
\item $s_i^2 = e$ for all $i$;
\item $(s_is_j)^{m_{ij}} = e$ for all $i \neq j$;
\item For any chordless cycle $C$ in $\Gamma$, where
\begin{displaymath}
C = i_0 \rightarrow i_1 \rightarrow \cdots \rightarrow i_{d-1} \rightarrow i_0
\end{displaymath}
and all of the weights $w_k$ are $1$ or $w_0 = 2$, we have
\begin{displaymath}
(s_{i_0}s_{i_1}\cdots s_{i_{d-2}}s_{i_{d-1}}s_{i_{d-2}}\cdots s_{i_1})^2 =e.
\end{displaymath}
\end{enumerate}

Using this group presentation, Barot and Marsh state the following result:
\begin{thm}\cite{BM13}[Theorem A]
Let $\Gamma$ be the diagram associated to a seed in a cluster algebra of finite type. Then $W(\Gamma)$ is isomorphic to the corresponding reflection group.
\end{thm}

@@

\begin{defn}
Let $W_{\Gamma}$ be the group with generators $s_i, i = 1,2,\ldots, n$, subject to the following relations:
\begin{itemize}
\item{(R1)} $s_i^2 = e$ for all $i$
\item{(R2)} $\left(s_is_j\right)^{m_{ij}} = e$ for all $i \neq j$
\end{itemize}
Furthermore, for a chordless cycle $C : i_0 \rightarrow i_1 \rightarrow \cdots \rightarrow i_{d-1} \rightarrow i_0$ and for $a = 0,1,2,\ldots, d-1$, define \textbf{$r\left(i_a, i_{a+1}\right) = s_{i_a}s_{i_{a+1}} \cdots s_{i_{a+d-1}}s_{i_{a+d-2}} \cdots s_{i_{a+1}}$}.\\

\vspace{0.1cm}
Then we have the following relations:
\begin{itemize}
\item{(R3)(a)} If all the weights in the edges of $C$ are $1$, then $r(i_a, i_{a+1})^2 = e$
\item{(R3)(b)} If $C$ has some edges of weight $2$, then $r(i_a, i_{a+1})^k = e$ where $k = 4-w_a$ and $w_a$ is the weight of the edge $i_a\--\ i_{a-1}$
\end{itemize}
\end{defn}

Defining the group $W_{\Gamma}$ with relations as shown above allows them to prove certain characteristics about the interaction between the relations in this group for the chordless cycles underlying the diagrams in question. In particular, they prove the following result, which we include as our main goal was to define analogous relations for an Artin group that would satisfy the same result.
\begin{thm}\cite{BM13}[Theorem 5.4a]
Let $\Gamma$ be a diagram of finite type and $\Gamma^{\prime} = \mu_k(\Gamma)$ the mutation of $\Gamma$ at vertex $k$. Then $W_{\Gamma} \cong W_{\Gamma^{\prime}}$.
\end{thm}

@@

\section{Proof of Main Result}\label{sec:proof_of_main}
